---
title: "Lim_Project_Task1"
author: "sheily lim"
date: "10/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```


## Importing data set

Big question: Did US intervention beginning in 2001 help improve Afghanistan's
human rights record?

```{r}
human_rights <- read.csv("Data/human-rights-scores.csv")
head(human_rights)


library(dplyr)
library(tidyverse)
library(ggplot2)
library(stargazer)
# install.packages("stargazer")
# install.packages("broom")
library(broom)

# The higher the score, the better. Ranges from -3.8 to 5.4
```

## Cleaning the data (human rights)

```{r}
# renaming score column
human_rights <- rename(human_rights, Score = Human.Rights.Score..Schnakenberg...Fariss..2014..Fariss..2019.)

human_rights<- human_rights %>% filter(Code == "AFG")
```

## Some statistical analysis (human rights)
This shows the overall mean human rights score. It shows when the min and max occured, as well
as what the mean was over pre and post US intervention
We observe the mean is lower under US intervention

```{r}

# records the max and min
max_rights <- max(human_rights$Score)
min_rights <- min(human_rights$Score)

# creates a column that shows time passed since 1946 for easier regression fitting
human_rights$years_passed <- human_rights$Year - 1946

#creates a column that has 1 for post US intervention, 0 for pre
human_rights$pre_post <- as.factor(ifelse(human_rights$Year >= 2001, 1, 0))

# finds mean of pre-US HR score
pre_us_mean_hr <- mean(filter(human_rights, pre_post == 0)$Score)

# finds mean of post-US HR score
post_us_mean_hr <- mean(filter(human_rights, pre_post == 1)$Score)

# locates the year where min and max HR score occur 
human_rights[human_rights$Score == max_rights,]
human_rights[human_rights$Score == min_rights,]

# ttest for statistical significance
ttest <- t.test(filter(human_rights, pre_post == 0)$Score, filter(human_rights, pre_post == 1)$Score)
tidy(ttest)
```


Here we are plotting the human rights score over time
We see there is a steep drop in 1978 from -.91 to -2.92
```{r}
# plot scores over time
hr_scatter <- ggplot(dat = human_rights, aes(x = Year, y = Score)) +
  geom_point(aes(color = pre_post)) +
  geom_line(aes(color = pre_post)) +
  geom_smooth(method = "lm", aes(color = pre_post), se = F) +
  geom_smooth(method = "lm", se = F) +
  labs(title = "Human Rights Scores in Afghanistan Over Time") +
  theme_classic() 

hr_scatter

# check the table around the years where we see the steep drop, identify it (1978)
human_rights[human_rights$Year > 1976 & human_rights$Year < 1983,]
```

Running regressions for overall HR score and years passed, then separated by pre
and post intervention 
```{r}

# relationship between human rights score and number of years passed 
lm_total <- lm(human_rights$Score ~ human_rights$years_passed)

# relationship between human rights score and years passed before intervention
lm_pre <- lm(filter(human_rights, pre_post == 0)$Score ~ filter(human_rights, pre_post == 0)$years_passed)

# relationship between human rights score and years passed after intervention
lm_post <- lm(filter(human_rights, pre_post == 1)$Score ~ filter(human_rights, pre_post == 1)$years_passed)

# print out table with all regression values
stargazer(lm_total, lm_pre, lm_post, type = "html", dep.var.labels = c("Total Time", "Pre-Intervention", "Post Intervention"), title = "Linear Regressions", out = "hr_regression1.htm")

stargazer(human_rights, type = "html", title = "Human Rights Scores", out = "hr_summary.htm")


```

Visualization: Bin the HR scores and then barplot
```{r}
# creating bins for the HR scores
breaks = c(-4, -3, -2, -1, 0, 1, 2)
labels = c("-3", "-2", "-1", "0", "1", "2")

# adding a column to the df with the binned data
human_rights$binned_scores <- cut(human_rights$Score, breaks = breaks, labels = labels)


hr_barplot <- ggplot(dat = human_rights, aes(x = binned_scores, color = (pre_post))) +
  geom_bar() +
  labs(x = "Human Rights Score", y = "Count", title = "Count of Human Rights Scores in Afghanistan since 1946") + 
  theme_classic()
```